---
title: "Homework-4-Functions-APIs"
format: pdf
editor: visual
---

## Task 1: Conceptual Questions
### Question 1
lapply() applies a function to each element of a vector or list. The equivalent purrr function is map().

### Question 2
```{r}
#| eval: false
# lapply with anonymous function
lapply(my_list, FUN = function(df) {
                  cor(df, method = "kendall")
                })
```

### Question 3
purrr functions give a shorthand way to make anonymous functions. purrr gives a bit cleaner / more consistent way to apply functions to objects, especially when using tidyverse (it includes lots of additional helper functions).

### Question 4
A side-effect function modifies something outside the function, but doesn't necessarily automatically return the data as well.

### Question 5
When you call a function, it creates temporary function environments. R uses lexical scoping, so when you create a new variable like 'sd' in a function, it doesn't overwrite the sd function.

## Task 2: Writing R Functions
### Question 1
```{r}
# calculates and returns RMSE
getRMSE <- function(resps, preds, ...) {
  squared_errors <- (resps - preds)^2
  return(sqrt(mean(squared_errors, ...)))
}
```

### Question 2
```{r}
# setup
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
 
# call function
getRMSE(resp, pred)
 
# set 2 values to NA
resp[[1]] <- NA_real_
resp[[2]] <- NA_real_
# redo predictions based on NA vals being included
pred <- predict(lm(resp ~ x), data.frame(x))
 
# call function with and without specifying whether to disregard NA vals
getRMSE(resp, pred)
getRMSE(resp, pred, na.rm = TRUE)
```

### Question 3
```{r}
# calculates and returns MAE
getMAE <- function(resps, preds, ...) {
  absolute_errors <- abs(resps - preds)
  return(mean(absolute_errors, ...))
}
```

### Question 4
```{r}
# setup
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
 
# call function
getMAE(resp, pred)

# set 2 values to NA
resp[[1]] <- NA_real_
resp[[2]] <- NA_real_

# redo predictions based on NA vals being included
pred <- predict(lm(resp ~ x), data.frame(x))

# call function with and without specifying whether to disregard NA vals
getMAE(resp, pred)
getMAE(resp, pred, na.rm = TRUE)
```

### Question 5
```{r}
# wrapper function with default metrics param and ...
getMetrics <- function(resps, preds, metrics = c("RMSE", "MAE"), ...) {
  if (!(is.vector(resps) && is.numeric(resps) && is.vector(preds) 
        && is.numeric(preds))) {
    
    # warning message and return
    message("Both responses and predictions should be numeric vectors")
    return(invisible(NULL))
  }
  
  results <- list()
  
  # if RMSE in metrics param calc RMSE
  if ("RMSE" %in% metrics) {
    results$RMSE <- getRMSE(resps, preds, ...)
  }
  # if MAE in metrics param calc MAE
  if ("MAE" %in% metrics) {
    results$MAE <- getMAE(resps, preds, ...)
  }
  
  return(results)
}
```

### Question 6
```{r}
# setup
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

# call function
getMetrics(resp, pred, metrics = "RMSE")
getMetrics(resp, pred, metrics = "MAE")
getMetrics(resp, pred, metrics = c("RMSE", "MAE"))

# set 2 values to NA
resp[[1]] <- NA_real_
resp[[2]] <- NA_real_

# redo predictions based on NA vals being included
pred <- predict(lm(resp ~ x), data.frame(x))

# call function using all combos of metrics param and disregard NA vals
getMetrics(resp, pred, metrics = "RMSE", na.rm = TRUE)
getMetrics(resp, pred, metrics = "MAE", na.rm = TRUE)
getMetrics(resp, pred, metrics = c("RMSE", "MAE"), na.rm = TRUE)

# call function with incorrect data
getMetrics(data.frame(y = c(1)), data.frame(y_pred = c(2)))
```

## Task 3: Querying an API and a Tidy-Style Function
### Question 1
```{r}
#| warning: false
library(tidyverse)
library(httr)
library(jsonlite)

api_key <- "86074e5006bb450db86061bc67de258c"

# GET call with q, from, and apiKey query params
resp <- GET("https://newsapi.org/v2/everything", query = list(
  q = "Physics",
  from = "2025-05-24",
  apiKey = api_key
))
```

### Question 2
```{r}
# parse raw content to JSON and then to list
parsed <- fromJSON(rawToChar(resp$content))

# get articles column and convert to tibble
as_tibble(parsed$articles)
```

### Question 3
```{r}
# function that queries API for articles with subject, from, and api_key params
queryAPI <- function(subject, from, api_key) {
  resp <- GET("https://newsapi.org/v2/everything", query = list(
    q = subject,
    from = from,
    apiKey = api_key
  ))
  parsed <- fromJSON(rawToChar(resp$content))
  return(as_tibble(parsed$articles))
}

# call function
queryAPI("gamestop", "2025-05-30", api_key = api_key)
```